---
title: 'Milestone 1'
author: 'Anh Le, Eve Wicksteed'
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r libraries}
library(here)
library(tidyverse)
library(DT)
```


# Air Quality Data

## Introduction

Something about air quality...

## Data Description

We use this [dataset](https://archive.ics.uci.edu/ml/datasets/Air+Quality) from ...
The data has 9358 instances and 15 attributes.

## Exploring the dataset

```{r read data}
#airq <- readr::read_csv(here::here("data","AirQualityUCI.csv"))
airq <- readr::read_csv(here::here("data","airquality.csv"))
DT::datatable(airq)
```



### summary stats

look at missing values
range of values, 
what variables are available?

### graph 1

faceted plots of variables vs time

### graph 2

correlogram of pollutants

### graph 3

plot of temp, RH.

## Research question

## Plan of action

## References

S. De Vito, E. Massera, M. Piga, L. Martinotto, G. Di Francia, On field calibration of an electronic nose for benzene estimation in an urban pollution monitoring scenario, Sensors and Actuators B: Chemical, Volume 129, Issue 2, 22 February 2008, Pages 750-757, ISSN 0925-4005.


